Determinism, Complexity, and Predictability in Computer Performance

### 2014-08-01 response.tex

We have modified the introduction in order to define our goals more
clearly.

\emph{I think that the possibility to estimate the best prediction
  error in advance from a relatively simple WPE characteristic would
  be interesting and useful, if it could hold true for a wide class of
  systems. Such a question would deserve a careful study as that
  started in the presented work. However, I have a number of critical
  remarks and strongly doubt that the conjecture of the authors can be
  valid for a reasonably wide class of systems (not even speaking of
  ``all systems''). Other aspects of the paper are not so important,
  in my view, despite some of them are fresh and interesting
  (e.g. taking a computer as an object of prediction). My critical
  remarks are listed below.}

Computer dynamics spans the full range of behavior reported in the
nonlinear dynamics literature.  Standard nonlinear time-series
analysis of data from these systems has confirmed the existence of
fixed points, periodic orbits of different periods, chaotic
trajectories with various Lyapunov exponents, etc.  in their dynamics.
While it is certainly possible that the nonlinear dynamics of computer
performance is subtly different from the nonlinear dynamics of other
systems---in a manner that cannot be detected by traditional nonlinear
time-series analysis---we believe that our results will hold for time
series drawn from other physical systems.

We are certainly aware that this claim may not be true, however, and
we are in the process of the follow-up study suggested by the
reviewer.  We have found that the relationship posited in our paper
holds for a time series of voltages from a chaotic laser used in the
SFI prediction competition, for instance.  (It also holds for various
paradigmatic systems, such as the H\'{e}non map and the Lorenz system, as
discussed in a later section of this response.)  The image below is an
augmented version of our revised Figure 7 that includes all of these
examples.

@@

\smallskip

\emph{5) The benchmark system - a computer as an object of prediction
  - seems to be an unexpected and, probably, interesting choice but
  just as an additional illustration of a statement which should be
  first shown for well-controlled and well-understood paradigmatic
  systems (like low-dimensional maps or ordinary differential
  equations). Indeed, it is not known in advance what error one should
  expect for such a complicated object using a certain predictive
  technique, what technique is optimal or close to optimal, etc. If
  the main point of the research is the relationship between WPE and
  the best prediction error (if any), selection of such a complicated
  object as computer performance variations only introduces its own
  difficulties into the problem.}

Computers are indeed complicated objects, but we have studied their
dynamics extensively and established quite firmly that they are
garden-variety dynamical systems with surprisingly low-dimensional
dynamics [Mytkowicz et al. CHAOS, 2009].  Standard nonlinear
time-series analysis techniques on performance data measured from
running computers produce clean and consistent results, as mentioned
above: fixed points, bifurcations, chaotic attractors, and so on.

Complicated, unknown systems are where prediction really
matters---especially systems where a good prediction can make a
practical difference (e.g., reallocation of resources in a computer to
suit the dynamically changing needs of a program).  Well-known systems
are certainly good testcases.  As is also mentioned above, the
relationship posited in our paper does indeed hold for paradigmatic
systems like the H\'{e}non map and the Lorenz equations.  We had not
included these examples in our paper because they simply confirmed
well-known results, and because our focus was on confirming those
results in the context of noisy, real-world data.  \alert{[[Liz to do,
      then finish this wording: We have added a sentence in the
      conclusion (or maybe end of discussion) about the fact that the
      WPE vs. MASE relationship for these synthetic examples follows
      the empirically determined pattern in our paper.]]}

@@

\smallskip

\emph{7) As for the choice of the four prediction techniques, some of
  them seem superfluous. E.g. naive approach is just a global AR model
  of the zero order.  Random-walk is similar to the AR model of the
  first order where the coefficient is not estimated by the
  least-squares but set equal to 1 (or this is an ARIMA model with the
  first difference and ``zero ARMA part''). Inefficiencies of the
  ARIMA model as compared to those choices in the paper can represent
  an inappropriate procedure for the order selection rather than
  indicate inappropriateness of the entire ARIMA approach. Thus,
  linear ARIMA and nonlinear LMA might be sufficient (in combination
  with mathematical benchmark examples) to represent close-to-optimal
  prediction errors.}

It is certainly true that the naive and random-walk predictors can be
seen as subclasses of the ARIMA model family.  We have explicitly
included them, however, because they are used (sometimes to great
success) in practice and because that made for a more comprehensive exploration.

The {\tt auto.arima} method that was used to build all of the ARIMA
models in this paper---another ``off the shelf'' method in the
modeling and prediction communities---follows the procedure outlined
in Section V B to choose values for the model parameters.
Sometimes the choices made by that procedure (e.g., setting a
particular parameter to zero) produce a naive or random-walk
predictor.  Our method can help practioners know whether this is a
good or bad thing.  If a particular model performs poorly (compared to
what the WPE value might suggest), it may be an indication that the
chosen parameter values were suboptimal---or even that no member of
the ARIMA family will work and one should try a nonlinear model. A discussion to this effect has been added to Section
VI in hopes of clarifying this completely valid concern and to illustrate that this is a feature of our proposed heuristic.

@@

\noindent\emph{General remarks.}

\emph{I) Strict relationship between WPE and prediction error
  applicable to all systems is hardly possible. In particular, there
  is a notion of epsilon-entropy which tends to the KS entropy if
  epsilon (the cell size) tends to zero (see e.g.  [Boffetta et al.]
  cited above). For stochastic systems, the KS entropy is infinite and
  epsilon-entropy varies strongly with the scale epsilon. Seemingly,
  the permutation entropy estimate in this case will strongly depend
  on the choice of the word length L. The authors concentrate on L = 6
  and obtain coefficients of the regression between WPE and prediction
  error only for that choice and for their particular
  signals. However, other systems and other choices of L may easily
  lead to other regression coefficients so that deciding about the
  best possible prediction error from WPE seems impossible in
  general. It is good if such a decision is possible for a certain
  narrow class of systems. However, the authors do not even discuss to
  which class of systems their results apply.}

WPE values depend not only on wordlength, but also on \emph{data}
length---since one needs more data to get good statistics on the words
for the PE calculation.  These two interacting knobs are not easy to
set.  Perhaps as a result of this, wordlength choices are rarely
justified in the permutation entropy literature.

The extent of the discussion in [Bandt \& Pompe, {\sl PRL}, 2002], for
instance---the paper in which PE was proposed---is ``Our entropies are
calculated for different [$\ell$] but we do not to determine a limit
for large [$\ell$], although this is an interesting theoretical
problem. For practical purposes, we recommend 3\dots7...''  The
corresponding discussion in [Cao {\sl et al.}, {\sl PRE}, 2004] reads
``In their paper, Bandt \& Pompe recommended [$\ell$] of 3-7.  We
often found that 3 and 4 may still be too small, and a value of 5, 6,
or 7 seems to be the most suitable....''

@@

The right thing to do, as we argued in Section [[what]], is to compute
PE across a \emph{range} of wordlengths ($\ell$) and look for a flat
spot in the curve of PE versus $\ell$.  We used that method to choose
individual $\ell$ values for each time series studied here.  We
conjecture that a systematic approach like sill mitigate any variation
in the results, thereby preserving the shape of the WPE vs. MASE
curve.  We have not yet confirmed that conjecture, but we are in the
process of doing so.

Regarding generality: the goal of this paper was to explore model-free
heuristics for empirical time-series data, where one does not know
anything about the underlying system.  In situations like this,
equivalence classes are difficult to define.  We addressed
this---insofar as is possible in an experimental context---by choosing
a range of time-series data sets whose behavior spans a wide range of
dynamical behaviors.  For all of these examples---as well as for the
39 additional ones\footnote{H\'{e}non, Lorenz, SFI A, random walk,
  nonlinear transformations of the 30 [[svdtwo]] and [[svdsix]]
  signals} constructed for this rebuttal letter---the pattern in
Figure 7 remained virtually unchanged, as described in the following
entry.  This strongly suggests, but of course does not prove, that our
approach is general.  As the reviewer suggested, this will require
extensive study and evaluation in future work.

@@

PE is invariant under an invertible nonlinear transformation of the
values in the time series, but WPE is not.  But MASE varies \emph{in a
  complementary way}.  Ryan will do $x(t)$ and $5e^{x(t)}$ and show
that it's on the curve.

% Ryan will insert figure here with x(t) = svdone and svdsix
\begin{center}
    \includegraphics[width=0.8\columnwidth]{figs/nonlinear_transform_data}
\end{center}

This is consistent with our claim of generality; this is just another
time series.  A xform like this effectively changes the generating
process, but our model-free result handles it.

We have revised the Figure, and the associated text, to include all
120 points.  We have also moved from a linear fit to a log fit in this
curve to better match the theoretical relationship between MASE and
WPE, as mentioned above.  Finally, we now show a ``fuzzy'' version of
the fitted curve in an effort to emphasize that the fundamental claim
of the paper is not related to the crispness of the fit.

Furthermore, we have recomputed the fit using both the 120 computer performance
data time-series as well as the 35 nonlinear transforms and the H\'{e}non, Lorenz,
random walk, and SFI Data A time-series. Not only does the fit line hardly
change, but the error bounds are contained completely within those provided by
just the computer performance data. This can be seen in the following figure
where the blue fit and bounds are those from the computer performance traces and
the magenta ones are those including the additional 39 time-series.

\begin{center}
    \includegraphics[width=0.8\columnwidth]{figs/new_prediction_vs_entropy_extras_with_nonlinear_all_points}
\end{center}

### 2014-08-01 model.tex
removing an alert and fixing a HEAD issue
friday round I

MASE scores for all 360 experiments are tabulated in the four middle
columns in Table~\ref{tab:error}.
%\begin{table*}
%\caption{Mean absolute scaled error (MASE) scores and weighted
%  permutation entropies for all eight processes studied in this paper.
%  LMA $=$ Lorenz method of analogues; RW $=$ random-walk prediction.
%  \alert{please reorder the columns from left to right to match the
%    order in which the methods were introduced, rather than in the
%    reverse of that order}}
%  \begin{center}
%  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill} } cccccc}
%  \hline\hline Signal & LMA MASE & \arima MASE & na\"{i}ve MASE & RW MASE & WPE \\
%  \hline
%
%  \col       & $0.050 \pm 0.002$ & $0.599 \pm 0.211$ & $0.571 \pm 0.002$  & $1.001 \pm 0.002$ & $0.513 \pm 0.003$ \\
%
%  \gcc       & $1.530 \pm 0.021$ & $1.837 \pm 0.016$ & $1.797 \pm 0.010$  & $1.138 \pm 0.011$ & $0.943 \pm 0.001$ \\
%
%  \svdone    & $0.827 \pm 0.076$ & $0.714 \pm 0.075$ & $2.676 \pm 4.328$  & $0.933 \pm 0.095$ & $0.957 \pm 0.016$ \\
%
%  \svdtwo    & $1.279 \pm 0.020$ & $2.163 \pm 0.027$ & $3.054 \pm 0.040$  & $1.125 \pm 0.012$ & $0.846 \pm 0.004$ \\
%
%  \svdthree  & $0.619 \pm 0.021$ & $0.713 \pm 0.010$ & $31.386 \pm 0.282$ & $0.707 \pm 0.009$ & $0.716 \pm 0.006$ \\
%
%  \svdfour   & $0.779 \pm 0.036$ & $0.979 \pm 0.032$ & $2.661 \pm 0.074$  & $1.034 \pm 0.035$ & $0.825 \pm 0.008$ \\
%
%  \svdfive   & $0.718 \pm 0.048$ & $2.370 \pm 0.051$ & $20.870 \pm 0.192$ & $1.001 \pm 0.047$ & $0.678 \pm 0.007$ \\
%
%  \svdsix    & $0.739 \pm 0.068$ & $1.438 \pm 0.061$ & $2.197 \pm 0.083$  & $1.060 \pm 0.055$ & $0.748 \pm 0.011$ \\
%
%  \hline\hline
%  \end{tabular*}
%  \end{center}
% \label{tab:error}
%  \end{table*}%
%  
 \begin{table*}
\caption{Mean absolute scaled error (MASE) scores and weighted
  permutation entropies for all eight processes studied in this paper.
  LMA $=$ Lorenz method of analogues; RW $=$ random-walk prediction.
  \alert{please reorder the columns from left to right to match the
    order in which the methods were introduced, rather than in the
    reverse of that order}}
  \begin{center}
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill} } cccccc}
  \hline\hline 
Signal & RW MASE & na\"{i}ve MASE & \arima MASE & LMA MASE & WPE \\
\hline
  \col       & $1.001 \pm 0.002$ & $0.571 \pm 0.002$  & $0.599 \pm 0.211$ & $0.050 \pm 0.002$ & $0.513 \pm 0.003$ \\
  \gcc       & $1.138 \pm 0.011$ & $1.797 \pm 0.010$  & $1.837 \pm 0.016$ & $1.530 \pm 0.021$ & $0.943 \pm 0.001$ \\
  \svdone    & $0.933 \pm 0.095$ & $2.676 \pm 4.328$  & $0.714 \pm 0.075$ & $0.827 \pm 0.076$ & $0.957 \pm 0.016$ \\
  \svdtwo    & $1.125 \pm 0.012$ & $3.054 \pm 0.040$  & $2.163 \pm 0.027$ & $1.279 \pm 0.020$ & $0.846 \pm 0.004$ \\
  \svdthree  & $0.707 \pm 0.009$ & $31.386 \pm 0.282$ & $0.713 \pm 0.010$ & $0.619 \pm 0.021$ & $0.716 \pm 0.006$ \\
  \svdfour   & $1.034 \pm 0.035$ & $2.661 \pm 0.074$  & $0.979 \pm 0.032$ & $0.779 \pm 0.036$ & $0.825 \pm 0.008$ \\
  \svdfive   & $1.001 \pm 0.047$ & $20.870 \pm 0.192$ & $2.370 \pm 0.051$ & $0.718 \pm 0.048$ & $0.678 \pm 0.007$ \\
  \svdsix    & $1.060 \pm 0.055$ & $2.197 \pm 0.083$  & $1.438 \pm 0.061$ & $0.739 \pm 0.068$ & $0.748 \pm 0.011$ \\  
    \hline\hline
  \end{tabular*}
  \end{center}
 \label{tab:error}
  \end{table*}%

In view of the discussion at the end of the previous paragraph, the
fact that the values in the second-from-\alert{left} column are not
identically equal to 1.00 may be somewhat surprising.  This can happen
due to differences between in-sample and out-of-sample forecasting;
the last 10\% of \svdthree, for instance---the out-of-sample
signal---was more amenable to random-walk prediction, on the average,
than the first 90\%.  See Section~\ref{sec:results} for a deeper
discussion of this effect.

>>>>

MASE scores for all 360 experiments are tabulated in the four middle
columns in Table~\ref{tab:error}.
%\begin{table*}
%\caption{Mean absolute scaled error (MASE) scores and weighted
%  permutation entropies for all eight processes studied in this paper.
%  LMA $=$ Lorenz method of analogues; RW $=$ random-walk prediction.
%  \alert{please reorder the columns from left to right to match the
%    order in which the methods were introduced, rather than in the
%    reverse of that order}}
%  \begin{center}
%  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill} } cccccc}
%  \hline\hline Signal & LMA MASE & \arima MASE & na\"{i}ve MASE & RW MASE & WPE \\
%  \hline
%
%  \col       & $0.050 \pm 0.002$ & $0.599 \pm 0.211$ & $0.571 \pm 0.002$  & $1.001 \pm 0.002$ & $0.513 \pm 0.003$ \\
%
%  \gcc       & $1.530 \pm 0.021$ & $1.837 \pm 0.016$ & $1.797 \pm 0.010$  & $1.138 \pm 0.011$ & $0.943 \pm 0.001$ \\
%
%  \svdone    & $0.827 \pm 0.076$ & $0.714 \pm 0.075$ & $2.676 \pm 4.328$  & $0.933 \pm 0.095$ & $0.957 \pm 0.016$ \\
%
%  \svdtwo    & $1.279 \pm 0.020$ & $2.163 \pm 0.027$ & $3.054 \pm 0.040$  & $1.125 \pm 0.012$ & $0.846 \pm 0.004$ \\
%
%  \svdthree  & $0.619 \pm 0.021$ & $0.713 \pm 0.010$ & $31.386 \pm 0.282$ & $0.707 \pm 0.009$ & $0.716 \pm 0.006$ \\
%
%  \svdfour   & $0.779 \pm 0.036$ & $0.979 \pm 0.032$ & $2.661 \pm 0.074$  & $1.034 \pm 0.035$ & $0.825 \pm 0.008$ \\
%
%  \svdfive   & $0.718 \pm 0.048$ & $2.370 \pm 0.051$ & $20.870 \pm 0.192$ & $1.001 \pm 0.047$ & $0.678 \pm 0.007$ \\
%
%  \svdsix    & $0.739 \pm 0.068$ & $1.438 \pm 0.061$ & $2.197 \pm 0.083$  & $1.060 \pm 0.055$ & $0.748 \pm 0.011$ \\
%
%  \hline\hline
%  \end{tabular*}
%  \end{center}
% \label{tab:error}
%  \end{table*}%
%  
 \begin{table*}
\caption{Mean absolute scaled error (MASE) scores and weighted
  permutation entropies for all eight processes studied in this paper.
  LMA $=$ Lorenz method of analogues; RW $=$ random-walk prediction.
  }
  \begin{center}
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill} } cccccc}
  \hline\hline 
Signal & RW MASE & na\"{i}ve MASE & \arima MASE & LMA MASE & WPE \\
\hline
  \col       & $1.001 \pm 0.002$ & $0.571 \pm 0.002$  & $0.599 \pm 0.211$ & $0.050 \pm 0.002$ & $0.513 \pm 0.003$ \\
  \gcc       & $1.138 \pm 0.011$ & $1.797 \pm 0.010$  & $1.837 \pm 0.016$ & $1.530 \pm 0.021$ & $0.943 \pm 0.001$ \\
  \svdone    & $0.933 \pm 0.095$ & $2.676 \pm 4.328$  & $0.714 \pm 0.075$ & $0.827 \pm 0.076$ & $0.957 \pm 0.016$ \\
  \svdtwo    & $1.125 \pm 0.012$ & $3.054 \pm 0.040$  & $2.163 \pm 0.027$ & $1.279 \pm 0.020$ & $0.846 \pm 0.004$ \\
  \svdthree  & $0.707 \pm 0.009$ & $31.386 \pm 0.282$ & $0.713 \pm 0.010$ & $0.619 \pm 0.021$ & $0.716 \pm 0.006$ \\
  \svdfour   & $1.034 \pm 0.035$ & $2.661 \pm 0.074$  & $0.979 \pm 0.032$ & $0.779 \pm 0.036$ & $0.825 \pm 0.008$ \\
  \svdfive   & $1.001 \pm 0.047$ & $20.870 \pm 0.192$ & $2.370 \pm 0.051$ & $0.718 \pm 0.048$ & $0.678 \pm 0.007$ \\
  \svdsix    & $1.060 \pm 0.055$ & $2.197 \pm 0.083$  & $1.438 \pm 0.061$ & $0.739 \pm 0.068$ & $0.748 \pm 0.011$ \\  
    \hline\hline
  \end{tabular*}
  \end{center}
 \label{tab:error}
  \end{table*}%

In view of the discussion at the end of the previous paragraph, the
fact that the values in the second-from-left column are not
identically equal to 1.00 may be somewhat surprising.  This can happen
due to differences between in-sample and out-of-sample forecasting;
the last 10\% of \svdthree, for instance---the out-of-sample
signal---was more amenable to random-walk prediction, on the average,
than the first 90\%.  See Section~\ref{sec:results} for a deeper
discussion of this effect.

### 2014-08-01 results.tex
friday round I

The discussion in the previous paragraph brings to light an interesting feature of our proposed heuristic. The \arima algorithm that was used to build all of the ARIMA
models in this paper, follows the procedure outlined
in Section V B to choose values for the model parameters.
As discussed previously, this algorithm employs sophisticated tests, such as the KPSS test, to intelligently traverse and reduce model space and then uses well-known criterion for choosing the most-likely of the remaining models. However, it is quite possible that this procedure can choose a sub-optimal fit. Specifically, if the initial space of models being searched is not broad enough, or if one of the preliminary tests gives an erroneous result, then the best ARIMA model may be excluded or eliminated before likelihood is even examined. This circumstance is hard to detect. Maximal-likelihood  tests, such as AIC used in \arima, can only conclude which of the models provided is ``most likely"---but cannot conclude that \emph{none} of the models are likely. In contrast,
%Sometimes the choices made by that procedure produce a sub-optimal predictor, however there is nothing to say whether a different model from the ARIMA class would work better.
inappropriate MASE scores (compared to
what the WPE value might suggest) may suggest an inappropriateness in the order selection and parameter estimation procedure, which would flag to a practitioner to reevaluate the appropriateness of fit. For example,
%During the fitting process of \arima many subclasses of ARIMA models are eliminated based on various tests, such as the KPSS test. If these tests give erroneous results then entire subclasses of ARIMA models are never even considered. Our heuristic may be able to flag when this has occurred.
by examining the \arima MASE scores vs. WPE values of \col, we were able to see that, while all signal complexities were the same, some of the signals had much higher MASE values. This suggested to us inappropriate parameter selection for these time series. We then went back and manually assisted the automated fitting procedure, resulting in models with consistent MASE vs WPE score, and improved forecasts.

@@

The second effect that can skew MASE scores is nonstationarity.  Since
this metric is normalized by the error of a random-walk forecast
\emph{on the training signal}, differences between the test signal and
training signal can create issues.  This is why the MASE values in
Table~\ref{tab:error} are not identically one for every random-walk
forecast of every time series: the last 10\% of these signals is
significantly different from the first 90\%.  The deviation from 1.00
will depend on the process---whether it has multiple regimes, what
those regimes look like, and how it switches between them---as well as
the experimental setup (e.g., sensor precision and data length).
Nonetheless, this effect suggests another potentially useful
heuristic: if the MASE of a random-walk prediction of a time series is
significantly different from 1, this could be an indication that the
signal is nonstationary.  \alert{Should we point back to the obvious
  MASE wall in the figure to add weight to this?}  We are in the
process of exploring this idea.

Again, the curves in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} were determined from a finite set of
methods and data.  We put a lot of thought and effort into making
these experiments upon which these results are based both
representative and comprehensive.  The forecast methods involved range
from the simple to the sophisticated; the time-series data was sampled
from a system whose behavior spans the dynamical behavior space.
While we are cautiously optimistic about the generality of our
conclusions, more exploration will be required before we can make any
such claim.  Our preliminary work along those lines shows that data
from the Henon map \cite{henon}, the Lorenz system \cite{lorenz}, and
the SFI A data set \cite{sfi-data} all fall within the one-$\sigma$
volume of the fit in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} region, as do various nonlinear
transformations of \svdtwo, \svdfive and \svdsix.

>>>>

The discussion in the previous paragraph highlights the second finding
of this paper: the ability of our proposed heuristic to flag
inappropriate models.  \arima is an automated, mechanical procedure
for choosing parameters for an ARIMA model of a given data set.  While
the tests and criteria employed by this algorithm are sophisticated
(see Section~\ref{sec:arima}), the results can still be
sub-optimal---if the initial space of models being searched is not
broad enough, for instance, or if one of the preliminary tests gives
an erroneous result.  \arima \emph{always} returns a model, and it can
be very hard to detect when that model is bad.  Our results suggest a
way to do so: if the MASE score of an \arima model is out of line with
the WPE value, that can be an indication of inappropriateness in the
order selection and parameter estimation procedure.  By examining the
\arima MASE scores vs. WPE values of \col, we were able to see that,
while all signal complexities were the same, some of the signals had
much higher MASE values. This suggested to us inappropriate parameter
selection for these time series. We then went back and manually
assisted the automated fitting procedure, resulting in models with
consistent MASE vs WPE score, and improved forecasts.

% The discussion in the previous paragraph brings to light an
% interesting feature of our proposed heuristic. The \arima algorithm
% that was used to build all of the ARIMA models in this paper follows
% the procedure outlined in Section V B to choose values for the model
% parameters.  As discussed previously, this algorithm employs
% sophisticated tests, such as the KPSS test, to intelligently traverse
% and reduce model space and then uses well-known criterion for choosing
% the most-likely of the remaining models. However, it is quite possible
% that this procedure can choose a sub-optimal fit. Specifically, if the
% initial space of models being searched is not broad enough, or if one
% of the preliminary tests gives an erroneous result, then the best
% ARIMA model may be excluded or eliminated before likelihood is even
% examined. This circumstance is hard to detect. Maximal-likelihood
% tests, such as AIC used in \arima, can only conclude which of the
% models provided is ``most likely"---but cannot conclude that
% \emph{none} of the models are likely. In contrast,
% %Sometimes the choices made by that procedure produce a sub-optimal predictor, however there is nothing to say whether a different model from the ARIMA class would work better.
% inappropriate MASE scores (compared to what the WPE value might
% suggest) may suggest an inappropriateness in the order selection and
% parameter estimation procedure, which would flag to a practitioner to
% reevaluate the appropriateness of fit. For example,
% %During the fitting process of \arima many subclasses of ARIMA models are eliminated based on various tests, such as the KPSS test. If these tests give erroneous results then entire subclasses of ARIMA models are never even considered. Our heuristic may be able to flag when this has occurred.
% by examining the \arima MASE scores vs. WPE values of \col, we were
% able to see that, while all signal complexities were the same, some of
% the signals had much higher MASE values. This suggested to us
% inappropriate parameter selection for these time series. We then went
% back and manually assisted the automated fitting procedure, resulting
% in models with consistent MASE vs WPE score, and improved forecasts.

@@

The second effect that can skew MASE scores is nonstationarity.  Since
this metric is normalized by the error of a random-walk forecast
\emph{on the training signal}, differences between the test signal and
training signal can create issues.  This is why the MASE values in
Table~\ref{tab:error} are not identically one for every random-walk
forecast of every time series: the last 10\% of these signals is
significantly different from the first 90\%.  The deviation from 1.00
will depend on the process---whether it has multiple regimes, what
those regimes look like, and how it switches between them---as well as
the experimental setup (e.g., sensor precision and data length).  For
the processes studied here, these effects do not cause the MASE values
to exceed 1.15, but pathological situations (e.g., a huge switch in
scale right at the training/test signal boundary) could produce higher
values.  This suggests another potentially useful heuristic: if the
MASE of a random-walk prediction of a time series is significantly
different from 1, this could be an indication that the signal is
nonstationary.  We are in the process of exploring this idea.

Again, the curves in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} were determined from a finite set of
methods and data.  We put a lot of thought and effort into making
these experiments upon which these results are based both
representative and comprehensive.  The forecast methods involved range
from the simple to the sophisticated; the time-series data was sampled
from a system whose behavior spans the dynamical behavior space.
While we are cautiously optimistic about the generality of our
conclusions, more exploration will be required before we can make any
such claim.  Our preliminary work along those lines shows that data
from the Henon map \cite{henon}, the Lorenz system \cite{lorenz}, and
the SFI A data set \cite{weigend-book,sfi-data} all fall within the
one-$\sigma$ volume of the fit in Figures~\ref{fig:wpe_vs_mase_best}
and~\ref{fig:wpe_vs_mase_all} region, as do various nonlinear
transformations of \svdtwo, \svdfive and \svdsix.
